Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-23 paper/high-assurance-contracts.tex
re-re-org

>>>>

\subsection{Code contracts}
\label{code-contracts}

Generally, AGREE specifications do not describe the computation that a
component performs. This is entirely by design: AGREE is intended to
reason about component behavior solely at the specification
level. However, the syntax of AGREE specifications provides enough
expressiveness to support the notion of a \emph{code contract}: a
contract from which an implementation can be extracted. First we must
discuss a class of guarantees---\emph{output guarantees}---which
determine the values on all output ports of a component.

\begin{definition}[Output guarantee]
An \emph{output guarantee} is a stylized guarantee that fully
specifies the data written to an output port. There are three
possibilities according to whether the output port $p$ is
a \konst{data} port, an \konst{event} port, or an \konst{event data}
port:
\[
\begin{array}{ll}
\konst{data}: &  p = \mathit{e} \\
\konst{event}: &  \konst{event} (p) = \mathit{b} \\
\konst{event data}: & \itelse{b}{\konst{event} (p) \land p = e}{\neg \konst{event}(p)} \\
\end{array}
\]
\end{definition}

Informally, a code contract treats its \konst{eq} ``statements'' as
defining a list of assignments to state variables, and its output
guarantees as directives for producing output.

\begin{definition}[Code contract] A
  leaf component of the form $(I,O,A,P,\emptyset)$ is a
  \emph{code contract} if $\mathit{Eqs} \cup G \subseteq P$, where
\[\mathit{Eqs} = \set{v_1 = e_1, \cdots , v_n = e_n} \] is a non-empty set
of \konst{eq} statements and $G$ is the set of output guarantees, one for
each element of $O$. In the interpretation as code, the order of
elements of $\mathit{Eqs}$ is important, and is simply taken to be the
occurrence order of the \konst{eq} statements in the syntax. Thus we
will work with the
\emph{list} of equations $\mathit{Eqs} = [v_1 = e_1; \cdots ; v_n = e_n]$.
\end{definition}

The semantics of Section \ref{agree-semantics} supports a formal
connection between the original contract (and verification results of
the AGREE model checker being run on it), and code generated from
specifications of leaf level cyber-components. It also provides the
root meaning at the base of a chain of translation steps moving from
AGREE specifications to CakeML executables.

The first step in the chain maps the contract to a code-focused
representation. Assume given code contract $(I,O,A,\mathit{Eqs} \cup G \cup
P,\emptyset)$, environment $E$, and time $t$. The evaluation of $\mathit{Eqs}$
followed by the evaluation of $G$ results in a new environment $E'$
where the stream values for state variables and outputs have been
computed for time $t$. The step from $E$ to $E'$ is one full cycle in
the repeated evaluation of the component.

\begin{definition}[Evaluation]
We overload existing notation and write the evaluation of
$\mathit{Eqs}$ and then $G$ in environment $E$ at time $t$ as
$\sem{\mathit{Eqs}\cdot G}^E_t$. Each $v = e \in {\Eqs}$ is treated as
an environment transformation
\[
 \sem{v = e}^E_t = E[v(t) \mapsto\sem{e}^E_t] \ .
\]
modifying $E$ so that stream $v$ has value $\sem{e}^E_t$ at time $t$.
{\Eqs} is evaluated by folding the transformation left-to-right
through it. The transformation is also applied to the list of output
guarantees $G$, computing the values on output ports. The details are
omitted, being a bit messy because of the different kinds of output
ports available.
\end{definition}

\begin{definition}[Code contract correctness]
Assume code contract $(I,O,A,\mathit{Eqs} \cup G \cup P,\emptyset)$.
The contract is \emph{correct}, if for all $E$ and $t$, whenever the
assumptions (historically) hold in $E$ and evaluation steps from $E$
to $E'$, then the guarantees hold in $E'$:
\[
\sem{\konst{Hist}(A)}^E_t \land E' = \sem{\mathit{Eqs} \cdot G}^E_t \imp \sem{P}^{E'}_t
\]
\end{definition}
\footnote
{KLS: At this point we would like to assert that the AGREE ``consistency
 check'' for a contract implies code contract correctness.}

How does this notion of correctness integrate with the verification
conditions addressed by model checking? In \emph{contract
correctness}, the model checker is proving the following property
\[
\konst{Hist}(A) \imp P
\]
and in \emph{code contract correctness} we essentially prove a Hoare
triple of the form
\[
\set{\konst{Hist}(A)}\; (\mathit{Eqs} \cdot G) \; \set{P}
\]
This accords with intution, namely that AGREE is a kind of
`program-free program logic'; by pulling out a program $(\mathit{Eqs}
\cdot G)$ from a code contract we find ourselves in the tepid
bathwater of imperative program verification. In order to use the code
contract result in system contract verifcation conditions, one has to
show that the code contract result implies the system verification
condtion.

\subsubsection*{Wellformedness}
The evaluation semantics given above is based on the stream
computation model, where values `deep in the past' can be accessed in
computations. However, the conventional program languages we target
target do not support such a view. Thus code generation is only
defined for a class of \emph{wellformed} equations.

\begin{definition}[Wellformedness]
\end{definition}

\subsection{Code contracts for the hardened system}
